---
title: "Homework 4"
author: "Jordan Glendrange"
date: "3/7/2021"
output: html_document
---

## Read Data

The table is stored in a csv on my github accounts. Using "read_csv" to read data into a dataframe.

```{r setup, include=FALSE}
flights <- read.csv('https://raw.githubusercontent.com/jglendrange/DATA607/main/flights.csv')
head(flights)
```

## Narrow Dataset

It will be difficult to extract the flight counts when their stored in different columns. Here I use the gather function to narrow the data set and store all of the flight counts in one column.

```{r}
library(tidyr)
library(dplyr)

flights_narrow <- flights %>%
  gather(City, Flights, 3:7)

head(flights_narrow)
```

## Summarise 

To discover which airline is more punctional we use a couple functions from the dplyr package. Namely mutate, to single out on time flights, and summarize to calculate the on time rate.

```{r}
flights_narrow %>%
  mutate(on_time = ifelse(Status=='on time', Flights, 0)) %>%
  group_by(Airline) %>%
  summarise(
    on_time_rate = sum(on_time) / sum(Flights)
  )
```

Here we can see AM WEST is slightly better than Alaska w/ being on time. Let's investigate further and see if there are any stand out cities.

```{r}
flights_narrow %>%
  mutate(on_time = ifelse(Status=='on time', Flights, 0)) %>%
  group_by(City) %>%
  summarise(
    on_time_rate = sum(on_time) / sum(Flights)
  )
```

Interesting. San Francisco performs the worst in terms of on time flights.

```{r}
flights_narrow %>%
  mutate(sf_flights = ifelse(City=='San.Francisco', Flights, 0)) %>%
  group_by(Airline) %>%
  summarise(
    sf_flights = sum(sf_flights) / sum(Flights)
  )
```

A larger share of Alaska's flights are flying into SF. This could be the reason it has a lower proportion of flights being delayed. Not because it's an inferior Airline. 