---
title: "Homework 3"
author: "Jordan Glendrange"
date: "February 21, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
```

# Problem 1

Here I am bringing the data in using the rvest and tidyverse libraries. Html_table brings in all the table into a list. Since I only have one table I select the first index. The "[-1,]" is to remove the first row of data since is was not relevant. The data is pretty messy, so for the purpose of the problem I remove all the other columns and just select the "MAJOR" column. Then I use a regex to match to either data or statistics while ignoring casing. 

```{r}
content <- read_html("https://projects.fivethirtyeight.com/mid-levels/college-majors/index.html")

tables <- content %>% html_table(fill = TRUE)

majors <- tables[[1]][-1,]
majors <- majors %>% select(2)

data_statistics_majors <- filter(majors, grepl('(data|statistics)', MAJOR, ignore.case = TRUE))

head(data_statistics_majors)
```

# Problem 2

Here I used the regular expression \\"[a-z\s?]+\\" to extract each element from the string. But, I was still left with commented out quotations. I used the remove all to get ride of these.

```{r}
fruits <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

print(fruits)
```


```{r}
fruit_new <- str_extract_all(fruits,'\\"[a-z\\s?]+\\"')[[1]]

print(fruit_new)

fruit_new <- str_remove_all(fruit_new, '\"')

print(fruit_new)
```

# Problem 3

1. (.)\1\1

**Answer**: This would match strings that repeat the same character of length 3. For example: 333, hhh, lll, ???

2. "(.)(.)\2\1"

**Answer**: This would capture strings with 4 characters looking at the first 2 and the same 2 characters in reverse order. For example: "assa", "1221", "1pp1"

3. (..)\1

**Answer**: Similar to one the strings characters would be repeated 4 times. The capture group is 2 characters here. For example: aaaa, bbbb, pppp

4. "(.).\1.\1"

**Answer**: This would capture all strings that repeat at the 1st, 3rd, and 5th indicies. The 2nd and 4th index can be any character. For example: "ahapa", "b1b4b", "p5p6p"

5. "(.)(.)(.).*\3\2\1"

**Answer**: This captures all strings that end with the same 3 characters it started with but in reverse order. The string can be any length. For example: 95122335159 See how the string starts with 951 and ends with 159.

# Problem 4

1. (.).*\1
2. .*(..).*\1.*
3. .*(.).*\1.*\1.*
