---
title: "Homework1"
author: "Jordan Glendrange"
date: "February 6, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

The data set I chose is titled "How Unpopular is Donald Trump?" The link can be found here: https://projects.fivethirtyeight.com/trump-approval-ratings/

The dataset is hosted on my github account.

```{r}
approval_poll <- read.csv("https://raw.githubusercontent.com/jglendrange/DATA607/main/approval_polllist.csv", TRUE, ",")

head(approval_poll)
```

The column names are pretty clear and straighforward. I decided to drop some columns due to data redundancy and mostly null values.

```{r}
wanted_columns <- c("pollster","subgroup","startdate","enddate","grade","samplesize","population","adjusted_approve","adjusted_disapprove","poll_id","question_id","createddate")
approval_poll <- approval_poll[wanted_columns]
head(approval_poll)
```

I decided to take a look at the approval and disapproval rattings via a historgram

```{r}
hist(approval_poll$adjusted_approve)
hist(approval_poll$adjusted_disapprove)

```

You can see easily see that the disapproval histogram skews further to the right.

For fun I decided to plot the dissapproval rating against the approval rating. I expected to see a 1 to 1 relationship, but was suprised to see a few outliers. It would be an interesting follow up to see which pollsters reported those numbers.

```{r}
plot(approval_poll$adjusted_disapprove, approval_poll$adjusted_approve)

```

## Conclusion

After reviewing the data set in the article Donald Trump held an average approval rating of ~42%. The only point Trump had an approval rating higher than his disapproval rating was at the very begining of his presidenciy. Furthermore at the end of his presidency there was a 20% gap between his approval and disapproval rating.

In terms of follow ups I would like to investigate - It would be interesting to dig into what each pollster is reporting to try and tease out any biases. 