---
title: "Homework Week 9"
author: "Jordan Glendrange"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library("httr")
library("tidyverse")
library("jsonlite")
```

## Request API

After creating my dev account with the New York Times website I created an app with an api key. I made an API request using the GET function from the HTTR library. Printing the response gives us a status code of 200 which means it was successful.

```{r}
r <- GET("https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key=uUhvu5PYBLtlbDqzEsy9uo8DlG4O7iM3")
print(r)
```

## Parsing the Response

We extract the text content from the response and then convert the json text into a dataframe using the function fromJSON(). Then we take the results out and convert that into a dataframe.

```{r}
text_content <- content(r, "text")

json_content <- text_content %>% fromJSON

books <- data.frame(json_content$results$books)

glimpse(books)
```

## Top Books

```{r}
books %>%  
  ggplot(aes(x = reorder(title, rank) , y = rank)) +
  geom_bar(stat="identity") +
  labs(
    x = "", y = "",
    title = "Which hardback books are the most popular?"
  ) +
  coord_flip() 
```

## Weeks on the top

```{r}
books %>%  
  ggplot(aes(x = reorder(title, weeks_on_list) , y = weeks_on_list)) +
  geom_bar(stat="identity") +
  labs(
    x = "", y = "",
    title = "Which hardback books have been on the top 15 the longest?"
  ) +
  coord_flip() 
```