---
title: "Lab 1: Intro to R"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

In the plot below we observe a sharp decline in female baptisms from 1640 to 1660.This would require further investigation as to why there was such a large decrease in that time period.

```{r trend-girls}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point()
```


### Exercise 3

In the plot below we are looking at the boy baptism count as a % of total baptisms. What we can see in the scatter plot below is that there was a pretty close to equal number of baptisms for males and females, but there we always more boys being baptized. The proportion never falls below 50%.

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)

arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_point()
```


### Exercise 4

After taking a glimpse we can see three columns in the present dataset.

year: the year the birth took place
boys: the count of boys born that year
girls: the count of girls born that year

```{r dim-present}
head(present)
```
The dataset appear to contain the years ranging from 1940 - 2002.

```{r}
present %>%
  summarize(min = min(year), max = max(year))
```


### Exercise 5

After taking the average counts of the 2 datasets, the present table is dealing with counts 3 orders of magnitude larger than the arbuthnot table.

```{r count-compare}
present %>%
  summarize(avg_boy = mean(boys), avg_girl = mean(girls))
```
```{r}
arbuthnot %>%
  summarize(avg_boy = mean(boys), avg_girl = mean(girls))
```

### Exercise 6

Looking at the birthrates we observe the same trend. Boys are being born in a greater proportion, however there seems to be a downward trend over the years.

```{r plot-prop-boys-present}
present <- present %>%
  mutate(birth_total = boys + girls)

present <- present %>%
  mutate(boy_proportion = boys / birth_total)

ggplot(data = present, aes(x = year, y = boy_proportion)) + 
  geom_point()
```


### Exercise 7

1961 is the year with the highest birth total in this data set.

```{r}
print(present %>%
  arrange(desc(birth_total)))
```

